
---
title: "Cancer Trials Explorer (CTE) -- examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{explorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr, warn = FALSE) # for data wrangling 
library(DT) # use datatables to format outcomes
```

The CTE library pulls data from several sources:

- [ClinicalTrials.gov](https://clinicaltrials.gov)

- [Drugbank](https://go.drugbank.com)

- [Drug Repurposing Hub](https://www.broadinstitute.org/drug-repurposing-hub)

- [TTD: Therapeutic Target Database](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC99057/)

## Filtering clinical trials

ClinicalTrials.gov hosts data for hundreds of thousand clinical trials (CTs) targeting a wide variety of conditions. 

The goal of this library is to focus on CTs interventional (as opposed to observational) treating a cancer-related condition. 

Accordingly, I included only CTs with terms related to cancer (e.g., cancer, tumor, leukemia, melanoma, glaucoma) in the conditions targeted by their intervention. 

## Data access

To access the data you first need to load the library.  

```{r setup}
library(CTE)
```

Then, the dataset is stored in separate files available in _lazy_ mode (i.e., loaded when needed).

For example, to query the list of cancer-related clinical trials in the data you can type the following:

Clinical trial basic information 

```{r}
data(ct_info)
ct_info
```

Clinical trial list of associated interventions (e.g., drug treatments)

```{r}
data(ct_interv)
ct_interv
```

Clinical trial list of associated targeted conditions

```{r}
data(ct_cond)
ct_cond
```

Drug bank list of drugs 

```{r}
data(drugbank)
drugbank
```

Repurposing Hub list of drugs with functional annotations

```{r}
data(repurp)
repurp
```

Gene targets from TTD

```{r}
data(gene_targets)
gene_targets
```


## Querying the data 

The library also includes functions to match drugs by name across the different datasets, or to query a particular drug name among the CT interventions. 

## Clinical trials listing bortezomib as intervention

For example, below I show how to obtain a catalog of interventions in cancer-related clinical trials matched with the string `bortezomib`. 


```{r}
tbl <- find_intervention("bortezomib")
```

The function returns a list of interventions matched with the name `bortezomib`. Each intervention is associated with a unique clinical trial ID, title, status, date of registration, a detailed description of the study, and the number of enrolled subjects (either expected or actual) for `r dplyr::n_distinct(tbl$id)` clinical trials (out of `r nrow(tbl)` entries). 
 
I use the function `datatable` from the package `DT` to show the results  as HTML table. 

```{r}
DT::datatable(dplyr::select(tbl, -enrollment,-enrollment_type)
    , rownames = FALSE
    , filter = "top"
    , extensions = "Buttons"
    , plugins = "ellipsis"
    , options = list (
        dom = 'Bfrtip'
        , buttons =  c('copy', 'csv')
        , columnDefs = list(list(
                targets = c(2, 5)
                , render = JS("$.fn.dataTable.render.ellipsis( 125 )")
                )
            )
        )
    )
```
